\documentclass[fleqn,10pt]{letter}
\usepackage[utf8]{inputenc}
\usepackage{units}
\usepackage{makecell}
\usepackage{float}
\usepackage[T1]{fontenc}
\usepackage[a4paper, margin=0.88in]{geometry}

\begin{document}

\begin{flushright}
Dr Carmen Cabrera\\
University of Liverpool\\
Liverpool, UK\\
18th of December, 2025\\
\end{flushright}

\begin{flushleft}
To Dr Shikha Ranjha \\
Associate Editor \\
\textit{Nature Communications} \\
\end{flushleft} 

Dear Dr Ranjha,

Thank you for inviting a major revision of our manuscript, ``Sustained changes to urban mobility after COVID-19 amplified socio-economic inequalities in Latin America''. We appreciate the careful editorial assessment and the constructive reports from the three reviewers.

We have undertaken a substantial revision that addresses the key concerns raised regarding data reliability, potential confounding factors, modelling choices and the scope of the discussion. In particular, we have made changes in four key areas. First, we have strengthened data validation and reliability checks, including new comparisons of our Facebook data with WorldPop data, explicit reporting of the extent and spatial distribution of data imputation and additional robustness tests addressing the validity of our data processing. Second, we have extended the modelling framework to account for additional confounding variables as advised by the reviewers, and to evaluate alternative model specifications (including GAMs and alternative trend extraction methods), demonstrating that our findings are robust to modelling choices and assumptions. Third, we have expanded and sharpened the ``Discussion'' section, with clearer articulation of limitations, stronger links to urban resilience and socio-economic equity and more concrete, context-specific policy implications for Latin American cities. Fourth, we have improved the clarity and presentation of our manuscript and Supplementary Information, including revisions to most figures and captions.

A detailed point-by-point response to all reviewer comments is provided, documenting how each concern has been addressed. We believe these revisions substantially strengthen the manuscript and resolve the issues identified in the initial review. We hope the revised submission will now be suitable for further consideration at Nature Communications.

Yours sincerely,

Dr Carmen Cabrera\\
Lecturer in Geographic Data Science\\
University of Liverpool

\end{document}